SubsidyExplorer - Step 3: Model Parameterization

### 2020-06-29 scripts/04_model_parameterization.Rmd
updated model parameters

This script handles the preliminary data wrangling necessary to create the biological parameters for the bio-economic model underlying the SubsidyExplorer toolkit. It outputs two CSV files used by the SubsidyExplorer toolkit (located in `r here::here("WTOToolkit", "data")`): 
- model-parameters-global.csv
- model-parameters-regional.csv

@@

```{r}
# One estimate of global carrying capacity
# global_k <- sum(costello_2016_dat$k)/0.78
global_k <- global_msy*9.6

@@

```{r}
# Calculate r
global_r <- (global_msy/global_k)*(phi+1)^(1/phi)

# Add r to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "r", value = global_r, source = "Costello et al. 2016"))
```

@@

```{r}
# Estimation of starting biomass
biomass_vec <- seq(0, 0.5*global_k, length.out = 2000)
growth_approx <- (((phi+1)/phi)*global_r*biomass_vec*(1-((biomass_vec/global_k)^phi)))
biomass_pos <- which.min(abs(growth_approx - (global_h0+8000000)))
global_b0 <- biomass_vec[biomass_pos]

@@

```{r}
# Calculate stats by region
costello_2016_dat_regional <- costello_2016_dat %>%
  dplyr::filter(region_fao_split != "") %>%
  mutate(msy_new = (msy/0.78)/n_regions) %>%
  left_join(regions_tidy, by = c("region_fao_split" = "fao_region")) %>%
  group_by(region) %>%
  summarize(msy = sum(msy_new, na.rm = T),
            price = median(price, na.rm = T)) %>%
  arrange(region)

# Extract msy
regional_msy <- round(costello_2016_dat_regional$msy)
names(regional_msy) <- region_names

@@

We estimate regional carrying capacities again by assuming K to be 9.6 times MSY.

```{r}
# Extract carrying capacity
regional_k <- regional_msy*9.6

@@

```{r}
# Calculate r regionally
regional_r <- (regional_msy/regional_k)*(regional_phi+1)^(1/regional_phi)

@@

```{r}
#Estimate starting biomass
regional_b0 <- numeric(length(region_names))
regional_biomass_growth <- c(4.1e6, 8e6, -2e6, 2.3e6, 2.3e6, -2e6)

for(i in 1:length(regional_h0)){
  biomass_vec <- seq(0, regional_k[i], length.out = 10000)
  growth_approx <- (((regional_phi[i]+1)/regional_phi[i])*regional_r[i]*biomass_vec*(1-((biomass_vec/regional_k[i])^regional_phi[i])))
  biomass_pos <- which.min(abs(growth_approx - (regional_h0 + regional_biomass_growth[i])))
  regional_b0[i] <- biomass_vec[biomass_pos] 
}
names(regional_b0) <- region_names

@@

```{r}
# Group by fleet and calculate base year harvest and effort for the global fleet
regional_fleet <- vessel_list_out %>%
  group_by(region) %>%
  summarize(vessels = n_distinct(ssvid),
            fishing_KWh = sum(fishing_KWh_eez_fao_ter, na.rm = T),
            bad_subs = sum(bad_subs, na.rm = T)) %>%
  arrange(region)

@@

# Add price constant to parameter table
# global_model_parameters <- global_model_parameters %>%
#   bind_rows(tibble(parameter = "delta", value = global_delta, source = "Costello et al. 2016"))
```

@@

```{r}
# Regional price constant
regional_delta <- costello_regional_dat$catch/(costello_regional_dat$price^regional_epsilon)

# Add price constant to parameter table
# regional_model_parameters <- regional_model_parameters %>%
#   bind_rows(c("parameter" = "delta", "source" = "Costello et al. 2016", regional_delta))
```

@@

```{r}
# Calculate global base prices
global_base_price1 <- ((1/global_delta)^(1/global_epsilon))*(sum(global_fleet$catch)^(1/global_epsilon))

CPI_2012 <- 229.594  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)
CPI_2018 <- 251.107  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)

global_base_price2 <- global_price_wrangling$price * (CPI_2018/CPI_2012)
```

@@

We then can calculate the rate of subsidization in the base year using the subsidy estimates from Sumaila et al.

```{r}
# Calculate global base subsidies
global_base_subsidies <- global_fleet$bad_subs/global_fleet$fishing_KWh

# Add to parameter table
# global_model_parameters <- global_model_parameters %>%
#   bind_rows(tibble(parameter = "s0_managed", value = global_base_subsidies[1], source = "Sumaila et al. 2019"))
# global_model_parameters <- global_model_parameters %>%
#   bind_rows(tibble(parameter = "s0_oa", value = global_base_subsidies[2], source = "Sumaila et al. 2019"))
```

And now we are able to estimate the cost coefficients for each fleet.

```{r}
# Calculate cost coefficients 
global_cost_coeff <- ((global_base_price1*global_fleet$catch) + (global_base_subsidies*global_fleet$fishing_KWh))/(global_fleet$fishing_KWh^global_beta)

# Add to parameter table
# global_model_parameters <- global_model_parameters %>%
#   bind_rows(tibble(parameter = "alpha_managed", value = global_cost_coeff[1], source = "Calculated"))
# global_model_parameters <- global_model_parameters %>%
#   bind_rows(tibble(parameter = "alpha_oa", value = global_cost_coeff[2], source = "Calculated"))
```

@@

# Add to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "eta", value = global_eta, source = "Calculated"))
```

@@

As with the global analysis, we assume $\beta$ to be equal to 1.3 across all regions.

```{r}
# Define global beta
regional_beta <- rep(1.3, length = length(region_names))
names(regional_beta) <- region_names

@@

```{r}
# Calculate base price regionally
regional_catches_total <- regional_fleet_harvest %>%
  gather(region, catch, -1) %>%
  group_by(region) %>%
  summarize(catch = sum(catch)) %>%
  spread(region, catch)

regional_base_price1 <- ((1/regional_delta)^(1/regional_epsilon))*(unlist(regional_catches_total[1,])^(1/regional_epsilon))

@@

As well as the rate of subsidization by fleet and region in the base year

```{r}
# get subsidy rates regionally by fleet
# Get effort
regional_fleet_subs <- regional_fleet %>%
  dplyr::select(region, fleet, bad_subs) %>%
  spread(region, bad_subs)

subsidies_regional <- tibble(fleet = c("managed", "open_access"))

for(i in 2:ncol(regional_fleet_subs)){

  out <- tibble(unlist(regional_fleet_subs[,i])/unlist(regional_fleet_effort[,i]))
  colnames(out) <- region_names[i-1]

  subsidies_regional <- subsidies_regional %>%
    bind_cols(out)

}

# Add subsidy base to parameter table
# regional_model_parameters <- regional_model_parameters %>%
#   bind_rows(c("parameter" = "s0_managed", "source" = "Sumaila et al. 2019; GFW 2019", unlist(subsidies_regional[1,-1])))
# regional_model_parameters <- regional_model_parameters %>%
#   bind_rows(c("parameter" = "s0_oa", "source" = "Sumaila et al. 2019; GFW 2019", unlist(subsidies_regional[2,-1])))
```

So we estimate cost coefficients for each fleet

```{r}

cost_coeff_regional <- tibble(fleet = c("managed", "open_access"))

for(i in 2:ncol(regional_fleet_subs)){

  out <- tibble(
    ((regional_base_price1[i-1]*unlist(regional_fleet_harvest[,i])) + 
      (unlist(subsidies_regional[,i])*unlist(regional_fleet_effort[,i]))) /
      (unlist(regional_fleet_effort[,i])^regional_beta[i-1])
    )
  colnames(out) <- region_names[i-1]

  cost_coeff_regional <- cost_coeff_regional %>%
    bind_cols(out)

}

# Add subsidy base to parameter table
# regional_model_parameters <- regional_model_parameters %>%
#   bind_rows(c("parameter" = "alpha_managed", "source" = "Calculated", unlist(cost_coeff_regional[1,-1])))
# regional_model_parameters <- regional_model_parameters %>%
#   bind_rows(c("parameter" = "alpha_oa", "source" = "Calculated", unlist(cost_coeff_regional[2,-1])))

And we again assume $\eta$ to be equal to 0.1 across all regions

@@

```{r}
# Save both parameter files
write_csv(global_model_parameters, paste0(results_dir, "global-model-parameters.csv"))
write_csv(global_model_parameters, paste0(app_dir, "global-model-parameters.csv"))

write_csv(regional_model_parameters, paste0(results_dir, "regional-model-parameters.csv"))
write_csv(regional_model_parameters, paste0(app_dir, "regional-model-parameters.csv"))

# Save vessel list
write_csv(vessel_list_out, paste0(results_dir, "vessel-list-2018-regions.csv"))
```

>>>>

This script handles the preliminary data wrangling necessary to create the biological parameters for the bio-economic model underlying the SubsidyExplorer toolkit. It outputs two CSV files used by the SubsidyExplorer toolkit (located in `r here::here("WTOToolkit", "data")`): 
- model_parameters_global.csv
- model_parameters_regional.csv

@@

```{r}
# One estimate of global carrying capacity
global_k <- sum(costello_2016_dat$k)/0.78
#global_k <- global_msy*9.1

@@

```{r}
# Calculate population growth rate
global_g <- (global_msy/global_k)*(phi+1)^(1/phi)

# Add g to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "g", value = global_g, source = "Costello et al. 2016"))

# Calculate intrinsic growth rate
global_r <- ((phi+1)/phi)*global_g

# Add g to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "r", value = global_g, source = "Costello et al. 2016"))
```

@@

```{r}
# Estimation of starting biomass
biomass_vec <- seq(0, 0.5*global_k, length.out = 10000)
growth_approx <- (((phi+1)/phi)*global_g*biomass_vec*(1-((biomass_vec/global_k)^phi)))
biomass_pos <- which.min(abs(growth_approx - (global_h0+8000000)))
global_b0 <- biomass_vec[biomass_pos]

@@

```{r}
# Calculate stats by region
costello_2016_dat_regional <- costello_2016_dat %>%
  dplyr::filter(region_fao_split != "") %>%
  mutate(msy_new = msy/n_regions,
         k_new = k/n_regions,
         catch_new = catch/n_regions) %>%
  left_join(regions_tidy, by = c("region_fao_split" = "fao_region")) %>%
  group_by(region) %>%
  summarize(msy = sum(msy_new, na.rm = T),
            k = sum(k_new, na.rm = T),
            catch = sum(catch_new, na.rm = T),
            price = median(price, na.rm = T)) %>%
  arrange(region)

# Extract msy
regional_msy <- round(costello_2016_dat_regional$msy/0.78)
names(regional_msy) <- region_names

@@

We estimate regional carrying capacities again by assuming K to be 9.1 times MSY.

```{r}
# Extract carrying capacity
#regional_k <- regional_msy*9.1
regional_k <- round(costello_2016_dat_regional$k/0.78)
names(regional_k) <- region_names

@@

```{r}
# Calculate r regionally
regional_g <- (regional_msy/regional_k)*(regional_phi+1)^(1/regional_phi)

# Add r to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "g", "source" = "Costello et al. 2016", regional_g))

# Calculate r regionally
regional_r <- ((regional_phi+1)/(regional_phi))*regional_g

@@

```{r}
#Estimate starting biomass
#regional_b0 <- numeric(length(region_names))
# regional_biomass_growth <- c(1.33e6, 1.33e6, 1.33e6, 1.33e6, 1.33e6, 1.33e6)
# 
# for(i in 1:length(regional_h0)){
#   biomass_vec <- seq(0, regional_k[i], length.out = 10000)
#   growth_approx <- (((regional_phi[i]+1)/regional_phi[i])*regional_r[i]*biomass_vec*(1-((biomass_vec/regional_k[i])^regional_phi[i])))
#   biomass_pos <- which.min(abs(growth_approx - (regional_h0 + regional_biomass_growth[i])))
#   regional_b0[i] <- biomass_vec[biomass_pos] 
# }

regional_b0 <- regional_k*0.2
names(regional_b0) <- region_names

@@

```{r}
# Group by fleet and calculate base year harvest and effort for the global fleet
regional_fleet <- vessel_list_out %>%
  group_by(region) %>%
  summarize(vessels = n_distinct(ssvid),
            fishing_KWh = sum(fishing_KWh_eez_fao_ter, na.rm = T),
            bad_subs = sum(bad_subs, na.rm = T),
            catch = sum(catch, na.rm = T)) %>%
  arrange(region)

@@

# Add price constant to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "delta", value = global_delta, source = "Costello et al. 2016"))
```

@@

```{r}
# Regional price constant
regional_delta <- costello_2016_dat_regional$catch/(costello_2016_dat_regional$price^regional_epsilon)

# Add price constant to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "delta", "source" = "Costello et al. 2016", regional_delta))
```

@@

```{r}
# Calculate global base prices
global_base_price1 <- ((1/global_delta)^(1/global_epsilon))*(global_h0^(1/global_epsilon))

# CPI_2012 <- 229.594  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)
# CPI_2018 <- 251.107  #BLS CPI All Items (Annual Ave), (BLS Series ID: CUUR0000SA0)
# 
# global_base_price2 <- global_price_wrangling$price * (CPI_2018/CPI_2012)
```

@@

@@

# Add to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "eta", value = global_eta, source = "Calculated"))

We also define an omega which is equal to 0.001.

```{r}
# Define global eta
global_omega <- 0.001

# Add to parameter table
global_model_parameters <- global_model_parameters %>%
  bind_rows(tibble(parameter = "omega", value = global_omega, source = "Calculated"))

@@

As with the global analysis, we assume $\beta$ to be equal to 1.3 across all regions. 
```{r}
# Define global beta
regional_beta <- rep(1.3, length = length(region_names))
names(regional_beta) <- region_names

@@

```{r}
# Calculate base price regionally
regional_base_price1 <- ((1/regional_delta)^(1/regional_epsilon))*(regional_fleet$catch^(1/regional_epsilon))

@@

And we again assume $\eta$ to be equal to 0.1 and $\omega$ to be equal to 0.001 across all regions

@@

```{r}
regional_omega <- rep(0.0011, length = length(region_names))
names(regional_omega) <- region_names

# Add subsidy base to parameter table
regional_model_parameters <- regional_model_parameters %>%
  bind_rows(c("parameter" = "omega", "source" = "Costello et al. 2016", regional_omega))
```

```{r}
# Save both parameter files
write_csv(global_model_parameters, paste0(results_dir, "model_parameters_global.csv"))
write_csv(global_model_parameters, paste0(app_dir, "model_parameters_global.csv"))

write_csv(regional_model_parameters, paste0(results_dir, "model_parameters_regional.csv"))
write_csv(regional_model_parameters, paste0(app_dir, "model_parameters_regional.csv"))
```
